---
title       : Timber volume prediction
subtitle    : Coursera Data Products course project
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

--- 

## Introduction

The application uses the _trees_ dataset (31 data points) to predict the volume of timber yielded by a black cherry tree as a function of it girth and height.

---


## Model


If we approximated a tree to a cylinder, we would have:
$$
\huge { Volume = Constant \times Height \times Girth^{2} }
$$
Taking the log:
$$
\huge { log(Volume) = Constant^{\prime} + log(Height) + 2 . log(Girth) }
$$

So we fit (non-reactively) a log-linear model of the volume, which gives us coefficients close to what we'd expect:

```{r}
data(trees)
fit <- lm(log(Volume) ~ log(Height) + log(Girth), data=trees )
fit$coefficients
```


---

## Reactive inputs

The application has two numeric inputs, where we enter the height and girth of the tree whose volume we want to predict.

For example, a height and girth of $\LARGE{76ft}$ and $\LARGE{12in}$ gives us a predicted volume of $\LARGE{ `r format(exp(predict(fit, data.frame(Girth=12, Height=76))), digits=4)  ` ft^3 }$.

---

## Results

The prediction is reactively calculated once:
```{r, eval=FALSE}
prediction <- reactive(exp(predict(fit, data.frame(Girth=input$girth, Height=input$height))))
```
and displayed on two plots (Volume ~ Girth and Volume ~ Height), in relation with the data.

```{r, echo=FALSE, fig.height=5, fig.align='center'}
par(mfrow=c(1 ,2))
pred <- exp(predict(fit, data.frame(Girth=12, Height=76)))

plot(Volume ~ Girth, data = trees, col='black', xlab="Girth (in)", ylab="Volume (cubic ft)")
points(12, pred, col='red', pch=19, cex=1.5)
legend("topleft", legend=c("Training Data", "Prediction"), pch=c(1, 19), col=c('black', 'red'))

plot(Volume ~ Height, data = trees, col='black', xlab="Height (ft)", ylab="Volume (cubic ft)")
points(76, pred, col='red', pch=19, cex=1.5)
legend("topleft", legend=c("Training Data", "Prediction"), pch=c(1, 19), col=c('black', 'red'))
``` 
